A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-09-15 report_clean_final/TEX/numerics.tex
fixed multiple typos extra

\section{Numerical Experiments}
This section is organized as follows. The first part is dedicated to the numerical implementation and simulation 
of the frequency domain formulation \eqref{vf1dcase}. We study the convergence of this formulation and the behaviour 
of the numerical solution as the absorption parameter $\nu$ tends to zero. The experiments in this section were 
performed with \urev{our} code written in Octave.  
We implemented the scheme described in Section \ref{sec:discr} for a case of $P_{1}$-space used 
for the approximation of $E_{x}$ and $E_{y}$ and $\theta=0$, thus working with System (\ref{eq:simple_system}). 
We apply permutation to the above system 
to obtain a 7-diagonal Hermitian matrix and solve the system with the Gauss back substitution algorithm. 
The intermediate part briefly validates the semi-lagrangian discretization of the time domain formulation.
The last part of the section deals with the question of the equivalence of the limiting absorption and limited amplitude 
principle. We compare the solutions obtained as $\nu\rightarrow 0$ with the help of \mrev{the frequency   
and time domain codes }(computed for large values of time). 
The time-domain code implements the scheme described in \urev{Section \ref{sec:staggered_discretization}} and 
is written in Fortran.

\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of the Implementation}
\label{sec:airy_fd_validity}
To check the validity of the code, we first perform a numerical experiment in the following case:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=-\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}, \qquad \mrev{\nu=0, \qquad \theta=0.}
\end{align}
\urev{In this case $E_{y}$ that solves (\ref{eq:main_frequency_domain})  satisfies the Airy equation, see also Remark \ref{remark:other},
\begin{align*}
 E_y''-xE_y=0,\qquad x\in [-L,\; H].
\end{align*}
The physical solution of this equation, the Airy function \urev{$\Ai(x)$}, see \cite[Chapter 10.4]{abramowitz_stegun}, is analytic and 
decays rapidly as $x\rightarrow \infty$. }
\mrev{Hence, choosing $H$ sufficiently large (the actual value used in the experiments $H=10$, where $\left|\Ai'(10)\right|<1e-9$) and the boundary conditions
\begin{align}
\label{eq:bcs_airy}
\partial_{x}E_{y}(-L)+2iE_{y}(-L)=2i\Ai(-L)+\Ai'(-L),\qquad 
\partial_{x}E_{y}(H)=0\approx \Ai'(H),
\end{align}
the exact solutions are: $\Ai(x)$ for the unknown $E_{y}$, and $i\frac{\delta(x)}{\alpha(x)}\Ai(x)$ for $E_{x}$. \urev{Hence both functions are smooth}. 
The well-posedness of the respective variational formulation is due to Remark \ref{remark:other}. }

In Fig.~\ref{fig:conv_rate} we verify the convergence rates for this problem, comparing the solution with the analytic solutions. 
Importantly, the obtained convergence rates are in agreement with known estimates of the standard theory of convergence~\cite[Chapter 5.7,
Chapter 0.4]{brenner}, \mrev{provided that the Aubin-Nitsche lemma \cite[Theorems 3.2.4,3.2.5]{ciarlet_fem} holds for the variational formulation.}  
\begin{figure}
\begin{tikzpicture}
    \begin{loglogaxis}[
        xlabel=$h$,
        ylabel=$Error$,
        width=0.4\textwidth,
        xmin=2e-6,
        ymin=1e-8,
        xmax=2,
	legend cell align=left,
        legend style={
at={(1.7,0.5)},
%legend pos=outer north east, 
anchor=east,
%legend pos=south east,
%anchor=east, 
draw=none,
font=\small
}
]\addplot[mark=*,black] table {pics_frequency_domain/data_airy/E1L2.dat}; 
    \addlegendentry{$\|E_{x}-E_{x}^{c}\|_{L^2}$};
    \addplot[mark=*,red, mark size=1pt] table {pics_frequency_domain/data_airy/E2H1.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{H^{1}}$};

          \addplot[mark=*,cyan, mark size=1pt] table {pics_frequency_domain/data_airy/E2L2.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{L^{2}}$};  
        \addplot[dotted] table{pics_frequency_domain/data_airy/E_h2.dat};
         \addlegendentry{$O(h^2)$};
         \addplot[dashed] table{pics_frequency_domain/data_airy/H.dat};
         \addlegendentry{$O(h)$};
    \end{loglogaxis}
    \end{tikzpicture}
    \caption{Convergence rates for the \mrev{problem described in Section \ref{sec:airy_fd_validity} (the problem (\ref{eq:main_frequency_domain}) discretized as (\ref{eq:simple_system})} 
    with the parameters (\ref{eq:parameters}) and the \mrev{boundary condition (\ref{eq:bcs_airy}))}. \mrev{Here $E_{y}^{c}(x)=\Ai(x)$ and 
    $E_{x}^{c}(x)=i\frac{\delta(x)}{\alpha(x)}\Ai(x)$.}}
    \label{fig:conv_rate}
\end{figure}

\subsubsection{Solution of the Frequency-Domain Problem with Resonance}
%\paragraph{Convergence}
Let us consider the case of the resonance, more precisely, we choose sufficiently smooth
$\alpha,\delta$, s.t. $\alpha(0)=0$ and $\delta(0)\neq 0$, and the solvability conditions 
of Lemma \ref{lemma:well_posedness}  are satisfied. 
For simplicity, let us assume \mrev{that $\alpha(x)$ is smooth and }
\begin{align}
\label{eq:cond}
 \mrev{\alpha(x)=-x \text{  in some neighbourhood of $0$ }}.
\end{align}
\mrev{Let $h>0$ denote the meshwidth. }Given $\mathbf{V}_{h}=P_{h}^{1}\times P_{h}^{1}$, with $P_{h}^{1}$ consisting of piecewise-linear (hat) functions, 
\mrev{and $E^{\nu}_{x},\;E^{\nu,h}_{x}$ being respectively the exact and the numerical solution, }
we look for the ratio $h(\nu)$ that would ensure the bound on the absolute error 
\begin{align}
\label{eq:problem1}
\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}<\epsilon,
\end{align}
given a fixed value of $\epsilon>0$ and $\nu\rightarrow 0$. Without loss of generality here we assume that $\nu>0$.

We use the following ingredients:
\begin{itemize}
 \item \lastrev{The C\'ea lemma} applied to the problem (\ref{vf1dcase}); here \mrev{$\mathbf{E}^{\nu}$ is the exact solution, 
 $\mathbf{E}^{\nu,h}$ is the computed solution}, and $C_c$ and $C_i$ are respectively the continuity and the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{\nu,h}\|_{\mathbf{V}}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in \mathbf{V}_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{\mathbf{V}}
 \leq C\nu^{-1}\min_{\mathbf{v}\in \mathbf{V}_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{\mathbf{V}},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and \lastrev{is valid} for $\nu\rightarrow 0$.   
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L^2}\leq C_1h^2\left\|v''\right\|_{L^2},\; C_1>0,\\
 \|v-I^{h}v\|_{H^1}\leq C_2h\|v''\|_{L^2},\; C_2>0, 
 \end{split}
\end{align}
where $I^{h}$ is an interpolation operator onto $P_{h}^{1}$. 
\item \mrev{An explicit bound on $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ based on \cite{Despres_2014} and \cite[Proposition 1]{singular_solutions}. 
First of all, we 
will assume that the main contribution to $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where, 
for sufficiently small $\nu$, the following expansion is valid \cite[Proposition 1]{singular_solutions}
\begin{align}
\begin{split}
\label{eq:exact}
 E_x^{\nu}(x)=\tilde{E}_x^{\nu}+S_{1}^{\nu}(x),\\
 \tilde{E}_x^{\nu}=f_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+\frac{g_{\nu}(x)}{\rho\lastrev{_{\nu}}(x-X_{\nu})},
 \end{split}
\end{align}
where \lastrev{$f_{\nu},\; g_{\nu}, \; S_{1}^{\nu}$} are smooth functions, 
which, together with their derivatives, are bounded independently of $\nu$ for all sufficiently small values of $\nu$; 
$\rho_{\nu}$ is a smooth function s.t. $\rho_{\nu}(0)=0,$ with the derivative $\rho_{\nu}'(0)=1$, 
the point $X_{\nu}\approx i\nu c+O(\nu^2)$, for some $c\in\mathbb{R}_{*}$. 
}
\item \mrev{An explicit bound on $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ based on \cite{singular_solutions}. As before, we 
will assume that the main contribution to $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where,
for sufficiently small $\nu$, it holds \cite[Proposition 1]{singular_solutions}
\begin{align*}
E_{y}^{\nu}(x)=\tilde{E}_{y}^{\nu}(x)+S^{\nu}_2(x), \\
\tilde{E}_{y}^{\nu}(x)=(\alpha(x)+i\nu)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu}),%=(-x+i\nu)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu}),
\end{align*}
where $\rho_{\nu},X_{\nu}$ are is in (\ref{eq:exact}), 
and $p_{\nu}(x),\; S_{2}^{\nu}(x)$ are smooth functions, which, together with their derivatives, are bounded independently of $\nu$ 
for all sufficiently small values of $\nu$.}
\end{itemize}
\lastrev{Let us recall that we defined the function space  $\mathbf{V}=L^{2}(\Omega)\times H^{1}(\Omega)$. }Estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$. 
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L^2}+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L^2}.
\end{align}
\mrev{
First of all let us examine the behaviour the derivatives of $E_{y}^{\nu}$ in the vicinity of the resonance, 
in particular, the term $\tilde{E}_{y}^{\nu}$:
\begin{align}
\label{eq:eynuprime}
 \left(\tilde{E}_{y}^{\nu}\right)'&= \alpha'(x)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+(\alpha(x)+i\nu)p'_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})
 +\frac{(\alpha(x)+i\nu)p_{\nu}(x)\rho_{\nu}'(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}.
\end{align}
The last term in this expression is well-defined in $x=0$, even for $\nu=0$ (the limit can be obtained with the help of the L'Hôpital's rule 
and recalling that $\rho_{\nu}(x-X_{\nu})=x-X_{\nu}+O((x-X_{\nu})^2)$ in the vicinity of the resonance), and is uniformly bounded for sufficiently small $\nu$, together with its derivatives. 
The same holds for the second term. To examine the behaviour of $\left(\tilde{E}_{y}^{\nu}\right)''$, 
we study solely the singular first term of (\ref{eq:eynuprime})
\begin{align*}
 \left(\alpha'(x)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})\right)'=\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}+(\alpha'(x)p_{\nu}(x))'\log\rho_{\nu}(x-X_{\nu}).
\end{align*}
The term $\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}$ can be bounded, for all sufficiently small $|x|$, by 
\begin{align*}
 \left|\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}\right|\leq \frac{C}{|x-i\nu c|},\; C>0,
\end{align*}
where we used an explicit form of $\rho_{\nu}(x-X_{\nu})=x-X_{\nu}+O((x-X_{\nu})^2)$ valid for small $\nu$ and $|x|$, as well as 
the expansion $X_{\nu}=ic\nu+O(\nu^2)$.  For sufficiently small $\nu$, the term $(\alpha'(x)p_{\nu}(x))'\log\rho_{\nu}(x-X_{\nu})$ 
can be bounded by the above expression as well.

@@

To derive a similar bound on the second derivative of $E_{x}^{\nu}$, we assume that the main contribution to $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$
comes from the singular part of $\tilde{E}_{x}^{\nu}$, whose second derivative can be represented in the following form, 
for some functions $F_{\nu,k}(x),\; k=0,..,3,$ uniformly bounded in $x,\; \nu$ in the vicinity of $(x,\nu)=(0,0)$:
\begin{align}
\label{eq:ex_eq1}
 \frac{d^2\tilde{E}_x^{\nu}}{dx^2}=\sum\limits_{k=1}^{3}\frac{F_{\nu,k}(x)}{\left(\rho_{\nu}(x-X_{\nu})\right)^{k}}+F_{\nu,0}(x)\log\rho_{\nu}(x-X_{\nu}).
\end{align}
}
Then there exists $c>0$, s.t. for all sufficiently small $\nu$ 
\bealn
 \left\|\frac{d^2E_x^{\nu}}{dx^2}\right\|_{L^2}^{2}&\leq c\int\limits_{-L}^{H}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. Plugging this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-\frac{3}{2}}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

>>>>

\section{Numerical Experiments}
This section is organized as follows. The first part is dedicated to the numerical implementation and simulation 
of the frequency domain formulation \eqref{vf1dcase}. We study the convergence of this formulation and the behaviour 
of the numerical solution as the absorption parameter $\nu$ tends to zero. The experiments in this section were 
performed with our code written in Octave.  
We implemented the scheme described in Section \ref{sec:discr} for a case of $P_{1}$-space used 
for the approximation of $E_{x}$ and $E_{y}$ and $\theta=0$, thus working with System (\ref{eq:simple_system}). 
We apply permutation to the above system 
to obtain a 7-diagonal Hermitian matrix and solve the system with the Gauss back substitution algorithm. 
The intermediate part briefly validates the semi-lagrangian discretization of the time domain formulation.
The last part of the section deals with the question of the equivalence of the limiting absorption and limited amplitude 
principle. We compare the solutions obtained as $\nu\rightarrow 0$ with the help of the frequency   
and time domain codes (computed for large values of time). 
The time-domain code implements the scheme described in Section \ref{sec:staggered_discretization} and 
is written in Fortran.

\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of the Implementation}
\label{sec:airy_fd_validity}
To check the validity of the code, we first perform a numerical experiment in the following case:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=-\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}, \qquad \nu=0, \qquad \theta=0.
\end{align}
In this case $E_{y}$ that solves (\ref{eq:main_frequency_domain})  satisfies the Airy equation, see also Remark \ref{remark:other},
\begin{align*}
 E_y''-xE_y=0,\qquad x\in [-L,\; H].
\end{align*}
The physical solution of this equation, the Airy function $\Ai(x)$, see \cite[Chapter 10.4]{abramowitz_stegun}, is analytic and 
decays rapidly as $x\rightarrow \infty$. 
Hence, choosing $H$ sufficiently large (the actual value used in the experiments $H=10$, where $\left|\Ai'(10)\right|<1e-9$) and the boundary conditions
\begin{align}
\label{eq:bcs_airy}
\partial_{x}E_{y}(-L)+2iE_{y}(-L)=\Ai'(-L)+2i\Ai(-L),\qquad 
\partial_{x}E_{y}(H)=0\approx \Ai'(H),
\end{align}
the exact solutions are: $\Ai(x)$ for the unknown $E_{y}$, and $i\frac{\delta(x)}{\alpha(x)}\Ai(x)$ for $E_{x}$. Hence both functions are smooth. 
The well-posedness of the respective variational formulation is due to Remark \ref{remark:other}.

In Fig.~\ref{fig:conv_rate} we verify the convergence rates for this problem, comparing the solution with the analytic solutions. 
Importantly, the obtained convergence rates are in agreement with known estimates of the standard theory of convergence~\cite[Chapter 5.7,
Chapter 0.4]{brenner}, provided that the Aubin-Nitsche lemma~\cite[Theorems 3.2.4, 3.2.5]{ciarlet_fem} holds for the variational formulation.  
\begin{figure}
\begin{tikzpicture}
    \begin{loglogaxis}[
        xlabel=$h$,
        ylabel=$Error$,
        width=0.4\textwidth,
        xmin=2e-6,
        ymin=1e-8,
        xmax=2,
	legend cell align=left,
        legend style={
at={(1.7,0.5)},
%legend pos=outer north east, 
anchor=east,
%legend pos=south east,
%anchor=east, 
draw=none,
font=\small
}
]\addplot[mark=*,black] table {pics_frequency_domain/data_airy/E1L2.dat}; 
    \addlegendentry{$\|E_{x}-E_{x}^{c}\|_{L^2}$};
    \addplot[mark=*,red, mark size=1pt] table {pics_frequency_domain/data_airy/E2H1.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{H^{1}}$};

          \addplot[mark=*,cyan, mark size=1pt] table {pics_frequency_domain/data_airy/E2L2.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{L^{2}}$};  
        \addplot[dotted] table{pics_frequency_domain/data_airy/E_h2.dat};
         \addlegendentry{$O(h^2)$};
         \addplot[dashed] table{pics_frequency_domain/data_airy/H.dat};
         \addlegendentry{$O(h)$};
    \end{loglogaxis}
    \end{tikzpicture}
    \caption{Convergence rates for the problem described in Section \ref{sec:airy_fd_validity} (the problem (\ref{eq:main_frequency_domain}) discretized as (\ref{eq:simple_system})
    with the parameters (\ref{eq:parameters}) and the boundary condition (\ref{eq:bcs_airy})). Here $E_{y}^{c}(x)=\Ai(x)$ and 
    $E_{x}^{c}(x)=i\frac{\delta(x)}{\alpha(x)}\Ai(x)$.}
    \label{fig:conv_rate}
\end{figure}

\subsubsection{Solution of the Frequency-Domain Problem with Resonance}
%\paragraph{Convergence}
Let us consider the case of the resonance, more precisely, we choose sufficiently smooth
$\alpha,\delta$, so that $\alpha(0)=0$ and $\delta(0)\neq 0$, and the solvability conditions 
of Lemma \ref{lemma:well_posedness}  are satisfied. 
For simplicity, let us assume that $\alpha(x)$ is smooth and 
\begin{align}
\label{eq:cond}
\alpha(x)=-x \text{  in some neighbourhood of $0$ }.
\end{align}
Let $h>0$ denote the meshwidth. Given $\mathbf{V}_{h}=P_{h}^{1}\times P_{h}^{1}$, with $P_{h}^{1}$ consisting of piecewise-linear (hat) functions, 
and $E^{\nu}_{x},\;E^{\nu,h}_{x}$ being respectively the exact and the numerical solution, 
we look for the ratio $h(\nu)$ that would ensure the bound on the absolute error 
\begin{align}
\label{eq:problem1}
\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}<\epsilon,
\end{align}
given a fixed value of $\epsilon>0$ and $\nu\rightarrow 0$. Without loss of generality here we assume that $\nu>0$.

We use the following ingredients:
\begin{itemize}
 \item \lastrev{The C\'ea lemma} applied to the problem (\ref{vf1dcase}); here $\mathbf{E}^{\nu}$ is the exact solution, 
 $\mathbf{E}^{\nu,h}$ is the computed solution, and $C_c$ and $C_i$ are respectively the continuity and the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{\nu,h}\|_{\mathbf{V}}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in \mathbf{V}_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{\mathbf{V}}
 \leq C\nu^{-1}\min_{\mathbf{v}\in \mathbf{V}_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{\mathbf{V}},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and \lastrev{is valid} for $\nu\rightarrow 0$.   
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L^2}\leq C_1h^2\left\|v''\right\|_{L^2},\; C_1>0,\\
 \|v-I^{h}v\|_{H^1}\leq C_2h\|v''\|_{L^2},\; C_2>0, 
 \end{split}
\end{align}
where $I^{h}$ is an interpolation operator onto $P_{h}^{1}$. 
\item An explicit bound on $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ based on \cite{Despres_2014} and \cite[Proposition 1]{singular_solutions}. 
First of all, we 
will assume that the main contribution to $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where, 
for sufficiently small $\nu$, the following expansion is valid \cite[Proposition 1]{singular_solutions}
\begin{align}
\begin{split}
\label{eq:exact}
 E_x^{\nu}(x)=\tilde{E}_x^{\nu}+S_{1}^{\nu}(x),\\
 \tilde{E}_x^{\nu}=f_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+\frac{g_{\nu}(x)}{\rho\lastrev{_{\nu}}(x-X_{\nu})},
 \end{split}
\end{align}
where \lastrev{$f_{\nu},\; g_{\nu}, \; S_{1}^{\nu}$} are smooth functions, 
which, together with their derivatives, are bounded independently of $\nu$ for all sufficiently small values of $\nu$; 
$\rho_{\nu}$ is a smooth function s.t. $\rho_{\nu}(0)=0,$ with the derivative $\rho_{\nu}'(0)=1$, 
the point $X_{\nu}\approx i\nu c+O(\nu^2)$, for some $c\in\mathbb{R}\setminus\{0\}$.

\item An explicit bound on $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ based on \cite{singular_solutions}. As before, we 
will assume that the main contribution to $\|\left(E_{y}^{\nu}\right)''\|_{L^{2}(\Omega)}$ comes from the behaviour in the vicinity of the resonance, where,
for sufficiently small $\nu$, it holds \cite[Proposition 1]{singular_solutions}
\begin{align*}
E_{y}^{\nu}(x)=\tilde{E}_{y}^{\nu}(x)+S^{\nu}_2(x), \\
\tilde{E}_{y}^{\nu}(x)=(\alpha(x)+i\nu)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu}),%=(-x+i\nu)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu}),
\end{align*}
where $\rho_{\nu},X_{\nu}$ are is in (\ref{eq:exact}), 
and \lastrev{$p_{\nu},\; S_{2}^{\nu}$} are smooth functions, which, together with their derivatives, are bounded independently of $\nu$ 
for all sufficiently small values of $\nu$.
\end{itemize}
\lastrev{Let us recall that we defined the function space  $\mathbf{V}=L^{2}(\Omega)\times H^{1}(\Omega)$. }Estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$. 
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L^2}+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L^2}.
\end{align}
First of all let us examine the behaviour the derivatives of $E_{y}^{\nu}$ in the vicinity of the resonance, 
in particular, the term $\tilde{E}_{y}^{\nu}$:
\begin{align}
\label{eq:eynuprime}
 \left(\tilde{E}_{y}^{\nu}\right)'&= \alpha'(x)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})+(\alpha(x)+i\nu)p'_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})
 +\frac{(\alpha(x)+i\nu)p_{\nu}(x)\rho_{\nu}'(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}.
\end{align}
The last term in this expression is well-defined in $x=0$, even for $\nu=0$ (the limit can be obtained with the help of the L'Hôpital's rule 
and recalling that $\rho_{\nu}(x-X_{\nu})=x-X_{\nu}+O((x-X_{\nu})^2)$ in the vicinity of the resonance), and is uniformly bounded for sufficiently small $\nu$, together with its derivatives. 
\nrev{The second term is also well-defined in $x=0$}. To examine the behaviour of $\left(\tilde{E}_{y}^{\nu}\right)''$, \nrev {it is sufficient to 
we study solely the singular first term of (\ref{eq:eynuprime})}
\begin{align*}
 \left(\alpha'(x)p_{\nu}(x)\log\rho_{\nu}(x-X_{\nu})\right)'=\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}+(\alpha'(x)p_{\nu}(x))'\log\rho_{\nu}(x-X_{\nu}).
\end{align*}
The term $\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}$ can be bounded, for all sufficiently small $|x|$, by 
\begin{align*}
 \left|\frac{\alpha'(x)p_{\nu}(x)\rho'_{\nu}(x-X_{\nu})}{\rho_{\nu}(x-X_{\nu})}\right|\leq \frac{C}{|x-i\nu c|},\; C>0,
\end{align*}
where we used an explicit form of $\rho_{\nu}(x-X_{\nu})=x-X_{\nu}+O((x-X_{\nu})^2)$ valid for small $\nu$ and $|x|$, as well as 
the expansion $X_{\nu}=ic\nu+O(\nu^2)$.  For sufficiently small $\nu$, the term $(\alpha'(x)p_{\nu}(x))'\log\rho_{\nu}(x-X_{\nu})$ 
can be bounded by the above expression as well.

@@

To derive a similar bound on the second derivative of $E_{x}^{\nu}$, we assume that the main contribution to $\|\left(E_{x}^{\nu}\right)''\|_{L^{2}(\Omega)}$
comes from the singular part of $\tilde{E}_{x}^{\nu}$, whose second derivative can be represented in the following form, 
for some functions $F_{\nu,k}(x),\; k=0,..,3,$ uniformly bounded in $x,\; \nu$ in the vicinity of $(x,\nu)=(0,0)$:
\begin{align}
\label{eq:ex_eq1}
 \frac{d^2\tilde{E}_x^{\nu}}{dx^2}=\sum\limits_{k=1}^{3}\frac{F_{\nu,k}(x)}{\left(\rho_{\nu}(x-X_{\nu})\right)^{k}}+F_{\nu,0}(x)\log\rho_{\nu}(x-X_{\nu}).
\end{align}

Then there exists $c>0$, s.t. for all sufficiently small $\nu$ 
\bealn
 \lastrev{\left\|\frac{d^2E_x^{\nu}}{dx^2}\right\|_{L^2}^{2}}&\leq c\int\limits_{-L}^{H}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. Plugging this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-\frac{3}{2}}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

Let us check numerically whether this holds true. 
To do so, we conduct the  numerical experiment with parameters for Problem 
(\ref{eq:main_frequency_domain}) as given in Table \ref{tab:parameters}. Recall as well that $\theta=0$. 
\begin{table}[htb!]
\begin{tabular}{l|l}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{lr}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{lr}
0, & x\leq -10,\\
-4/30x-4/3,& -10<x\leq 5,\\
-2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance.}
\label{tab:parameters}
\end{table}
Denoting by $E_{x}^{\nu,h}$ the solution $E_{x}(x)$ computed on the mesh with width $h$ and 
 absorption $\nu$, and by $E_{x}^{\nu}$ the solution computed on a mesh multiple times finer (more precisely, the reference mesh
was typically 10-200 times finer, except when the computation was performed for very small values of $h$ when results were compared with results computed on twice finer meshes), let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{\nu,h'}-E_{x}^{\nu}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/h_nu.pdf}
&
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/res_sol.pdf}
\end{tabular}
\caption{We consider the problem (\ref{eq:main_frequency_domain}) with the parameters 
defined in Table \ref{tab:parameters}. In the left figure the dependence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated.  
In the right figure we show the computed solutions for this problem (for brevity denoted by $E_{x,y}^{\nu}$) for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that Estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for given values of $\epsilon$ and for a chosen range of $\nu>0$.
\begin{remark}
It can be shown that $\|E^{\nu}_{x}\|_{L^2}\leq \frac{C}{\sqrt{\nu}},\; C>0$, 
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L^2}}{\|E^{\nu}_{x}\|_{L^2}}\leq \epsilon
\eealn
 is ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{2}}$, $\beta_{\epsilon}>0$.
\end{remark}
